Preregistration analysis, Cost 2 study

### 2019-05-10 Cost2/Cost2_prereg.Rmd
implemented a distinction between accepted and completed trials in both experiments

``` {r Load Data, echo=FALSE}
### data loading and cleaning up
setwd("./data")
files <- dir(pattern = '_log.csv')
nSubjs <- length(files)

dataAll <- tibble(SubjID = files) %>%
  mutate(contents = map(SubjID, ~ read_csv(., col_types = cols())))  %>%
  mutate(Cost = substring(SubjID, 5, 8),
         Cost = case_when(Cost == "wait" ~ "Wait",
                          Cost == "cogT" ~ "Cognitive",
                          Cost == "phys" ~ "Physical",
                          Cost == "phea" ~ "Easy"),
         SubjID = as.integer(substring(SubjID, 0, 3))) %>%
  unnest() %>%
  mutate(Half = ifelse(Block < 4, "Half_1", "Half_2"),
         Cost = factor(Cost, levels = c("Physical", "Cognitive", "Wait", "Easy")),
         optimal = case_when(
          (Handling == 10 & Reward < 8) ~ 0,
          (Handling == 14 & Reward < 20) ~ 0,
          TRUE ~ 1
         ))

@@

``` {r 16.1.3. Earnings, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
# calculate the proportions and earnings per subject, while keeping group info
temp <- dataAll %>%
  filter(Choice < 2) %>%
  group_by(SubjID, Cost) %>%
  summarise(Earnings = sum(Reward[Choice == 1] / 100),
            pComplete = mean(Choice)) %>%
  mutate(pComplete2 = pComplete^2) %>%
  ungroup()

@@

# plot
(earnFitplot <- ggplot(data = temp, aes(pComplete, Earnings)) + 
         scale_color_manual(values = cols) + 
         geom_point(aes(fill = Cost), pch = 21, color = "black", size = 3, show.legend = T) + 
         stat_smooth(aes(y = Earnings), method = "lm", formula = y ~ x + I(x^2), color = "grey30") +
         labs(x = "Proportion Accepted", y = "Total Earnings (dollars)") +
         scale_fill_manual(values = cols) +
         theme(panel.grid.major = element_blank(), 
           panel.grid.minor = element_blank(), 
           panel.background = element_blank(), 
           axis.line = element_line(colour = "black"),
           text = element_text(size = 16)))

@@

``` {r 16.1.4., echo=FALSE}
# in order
# clean data, 
# calculate each individual's prop. accept,
# calculate t.tests with optimality as null mean,
# correct pvalues with FDR
temp <- dataAll %>%
  filter(Choice < 2) %>%
  group_by(SubjID, Handling, Reward, Cost) %>%
  summarise(pComplete = mean(Choice)) %>%
  mutate(optimal = case_when(
          (Handling == 10 & Reward < 8) ~ 0,
          (Handling == 14 & Reward < 20) ~ 0,
          TRUE ~ 1
        )) %>% 
  group_by(Handling, Reward, Cost) %>%
  summarise(meanAccept = mean(pComplete),
            sdAccept = sd(pComplete),
            optimal = unique(optimal),
            pvals = tryCatch(t.test(pComplete, mu = unique(optimal))$p.value, error = function(e) {1})) %>%
  mutate(FDR = p.adjust(pvals, method = "BY"),
         significant = FDR < 0.05)

@@

``` {r 16.2.1. Overall Proportion/Earnings, fig.align = "center", fig.width = 8, fig.height = 4, echo = FALSE}  
## ECDF of proportion completed per cost type
temp <- dataAll %>%
        filter(Choice < 2) %>%
        mutate(Cost = factor(Cost, levels = list("Physical", "Cognitive", "Wait", "Easy"))) %>%
        group_by(SubjID, Cost) %>%
        summarise(Earnings = sum(Reward[Choice == 1] / 100),
                  pComplete = mean(Choice))

## formal testing 
# proportion completed
propAov <- aov(pComplete ~ factor(Cost), data = temp)

propTests <- temp %>% 
  plyr::dlply("Cost", identity) %>%
  sapply("[[", "pComplete") %>%
  as.data.frame() %>%
  combn(2, simplify = F) %>%
  sapply(function(x) {
    c(contrast = paste(colnames(x), collapse = "vs"), 
      permPval = round(permute(x[, 1], x[, 2], simple = T), digits = 3),
      cohenD = round(cohenD(x[, 1], x[, 2]), digits = 3))
  }) %>% 
  t()

@@

# plots
p1 <- ggplot(aes(pComplete, color = Cost), data = temp) +
        stat_ecdf(lwd = 1.2, show.legend = F) +
        scale_color_manual(values = cols) +
        xlim(0, 1) +
        labs(y="ECDF", x = "Proportion Accepted") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(), 
              panel.background = element_blank(), 
              axis.line = element_line(colour = "black"),
              text = element_text(size = 16))

p2 <- ggplot(temp, aes(Cost, pComplete, fill = Cost)) + 
        geom_boxplot(show.legend = F) +
        scale_fill_manual(values = cols) +
        ylim(c(0, 1)) +
        labs(y = "Proportion Completed") +
        theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), 
            panel.background = element_blank(), 
            axis.line = element_line(colour = "black"),
            text = element_text(size = 16))

p3 <- ggplot(temp, aes(Cost, Earnings, fill = Cost)) + 
        geom_boxplot(show.legend = F) +
        scale_fill_manual(values = cols) +
        labs(y = "Total Earnings (dollars)") +
        theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), 
            panel.background = element_blank(), 
            axis.line = element_line(colour = "black"),
            text = element_text(size = 16))

@@

``` {r 16.2.2. rmANOVA, echo=FALSE,fig.align="center",fig.width=5,fig.height=4, results = 'asis'}
temp <- dataAll %>%
        filter(Choice < 2) %>%
        group_by(SubjID, Cost, Handling, Reward) %>%
        summarise(pComplete = mean(Choice))

## REPEATED MEASURES ANOVA
# I think I like this the most
prop.aov <- with(temp, aov(pComplete ~ factor(Cost) * Handling * Reward + 
                                 Error(SubjID / (Handling * Reward))))

@@

``` {r 16.2.2. rmANOVA plot, echo = FALSE, fig.align = "center", fig.width = 10, fig.height = 6}
# prep data to plot the p(accept) based on reward, handling time and cost type
trueProp <- dataAll %>%
        filter(Choice < 2) %>%
        group_by(SubjID, Cost, Handling, Reward, optimal) %>%
        summarise(pComplete = mean(Choice)) %>%
        group_by(Cost, Handling, Reward, optimal) %>%
        summarise(meanComplete = mean(pComplete),
                  SE = sd(pComplete) / sqrt(length(pComplete)))

@@

``` {r 16.3.1. RTs, echo=FALSE,fig.align="center",fig.width=6,fig.height=4}
# Proportion of forced travels per subject and cost type
propFails <- dataAll %>% 
  filter(Cost != "Wait") %>% 
  group_by(SubjID, Cost) %>% 
  summarise(propFails = mean(Choice == 2)) %>%
  group_by(Cost) %>% 
  summarise(meanQuit = mean(propFails))

# Reaction times ECDFs including offer and in-trial quits
(RTs <- dataAll %>%
  #mutate(RT = ifelse(Choice != 0, RT + 2, RT)) %>%
  filter(Choice < 2) %>% # change this to 1 to reproduce the original plot, since it would only plot the cdf of quits
  ggplot(aes(RT, color = Cost)) + 
    #stat_ecdf(lwd = 1.2, geom = "smooth") +
    geom_line(aes(y = 1 - ..y..), stat = 'ecdf', lwd = 1.5) +
    geom_vline(xintercept = 1, lty = 2) + # end of the grace period for physical trials
    scale_color_manual(values = cols) +
    scale_x_continuous(breaks = seq(14)) +
    labs(x = "Reaction Time", y = "Proportion Accepted") +
    theme(panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(colour = "black"),
      text = element_text(size = 16)))

>>>>

``` {r Load Data, echo=FALSE}
### data loading and cleaning up
# recent addition: turning forced travels to acceptances, as they reflect a preference for that trial
# rawChoice will be used to compute the # of mistakes
setwd("./data")
files <- dir(pattern = '_log.csv')
nSubjs <- length(files)

dataAll <- tibble(SubjID = files) %>%
  mutate(contents = map(SubjID, ~ read_csv(., col_types = cols())))  %>%
  mutate(Cost = substring(SubjID, 5, 8),
         Cost = case_when(Cost == "wait" ~ "Wait",
                          Cost == "cogT" ~ "Cognitive",
                          Cost == "phys" ~ "Physical",
                          Cost == "phea" ~ "Easy"),
         SubjID = as.integer(substring(SubjID, 0, 3))) %>%
  unnest() %>%
  mutate(rawChoice = Choice,
         Choice = ifelse(Choice == 2, 1, Choice),
         Half = ifelse(Block < 4, "Half_1", "Half_2"),
         Cost = factor(Cost, levels = c("Physical", "Cognitive", "Wait", "Easy")),
         optimal = case_when(
          (Handling == 10 & Reward < 8) ~ 0,
          (Handling == 14 & Reward < 20) ~ 0,
          TRUE ~ 1
         ))

@@

``` {r 16.1.3. Earnings, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
# calculate the proportions and earnings per subject, while keeping group info
temp <- dataAll %>%
  filter(rawChoice < 2) %>%
  group_by(SubjID, Cost) %>%
  summarise(Earnings = sum(Reward[Choice == 1] / 100),
            pComplete = mean(Choice)) %>%
  mutate(pComplete2 = pComplete^2) %>%
  ungroup()

@@

# plot
(earnFitplot <- ggplot(data = temp, aes(pComplete, Earnings)) + 
         scale_color_manual(values = cols) + 
         geom_point(aes(fill = Cost), pch = 21, color = "black", size = 3, show.legend = T) + 
         stat_smooth(aes(y = Earnings), method = "lm", formula = y ~ x + I(x^2), color = "grey30") +
         labs(x = "Proportion Completed", y = "Total Earnings (dollars)") +
         scale_fill_manual(values = cols) +
         theme(panel.grid.major = element_blank(), 
           panel.grid.minor = element_blank(), 
           panel.background = element_blank(), 
           axis.line = element_line(colour = "black"),
           text = element_text(size = 16)))

@@

``` {r 16.1.4., echo=FALSE}
# in order
# clean data, 
# calculate each individual's prop. accept,
# calculate t.tests with optimality as null mean,
# correct pvalues with FDR
temp <- dataAll %>%
  filter(Choice < 2) %>%
  group_by(SubjID, Handling, Reward, Cost) %>%
  summarise(pAccept = mean(Choice)) %>%
  mutate(optimal = case_when(
          (Handling == 10 & Reward < 8) ~ 0,
          (Handling == 14 & Reward < 20) ~ 0,
          TRUE ~ 1
        )) %>% 
  group_by(Handling, Reward, Cost) %>%
  summarise(meanAccept = mean(pAccept),
            sdAccept = sd(pAccept),
            optimal = unique(optimal),
            pvals = tryCatch(t.test(pAccept, mu = unique(optimal))$p.value, error = function(e) {1})) %>%
  mutate(FDR = p.adjust(pvals, method = "BY"),
         significant = FDR < 0.05)

@@

``` {r 16.2.1. Overall Proportion/Earnings, fig.align = "center", fig.width = 8, fig.height = 4, echo = FALSE}  
## ECDF of proportion completed per cost type
temp <- dataAll %>%
        filter(Choice < 2) %>%
        mutate(Cost = factor(Cost, levels = list("Physical", "Cognitive", "Wait", "Easy"))) %>%
        group_by(SubjID, Cost) %>%
        summarise(Earnings = sum(Reward[rawChoice == 1] / 100),
                  pAccept = mean(Choice))

## formal testing 
# proportion completed
propAov <- aov(pAccept ~ factor(Cost), data = temp)

propTests <- temp %>% 
  plyr::dlply("Cost", identity) %>%
  sapply("[[", "pAccept") %>%
  as.data.frame() %>%
  combn(2, simplify = F) %>%
  sapply(function(x) {
    c(contrast = paste(colnames(x), collapse = "vs"), 
      permPval = round(permute(x[, 1], x[, 2], simple = T), digits = 3),
      cohenD = round(cohenD(x[, 1], x[, 2]), digits = 3))
  }) %>% 
  t()

@@

## plots
# ECDF pAcceptd
p1 <- ggplot(aes(pAccept, color = Cost), data = temp) +
        stat_ecdf(lwd = 1.2, show.legend = F) +
        scale_color_manual(values = cols) +
        xlim(0, 1) +
        labs(y="ECDF", x = "Proportion Accepted") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(), 
              panel.background = element_blank(), 
              axis.line = element_line(colour = "black"),
              text = element_text(size = 16))

# pAccepted
p2 <- ggplot(temp, aes(Cost, pAccept, fill = Cost)) + 
        geom_boxplot(show.legend = F) +
        scale_fill_manual(values = cols) +
        ylim(c(0, 1)) +
        labs(y = "Proportion Accepted") +
        theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), 
            panel.background = element_blank(), 
            axis.line = element_line(colour = "black"),
            text = element_text(size = 16))

# earnings
p3 <- ggplot(temp, aes(Cost, Earnings, fill = Cost)) + 
        geom_boxplot(show.legend = F) +
        scale_fill_manual(values = cols) +
        labs(y = "Total Earnings (dollars)") +
        theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), 
            panel.background = element_blank(), 
            axis.line = element_line(colour = "black"),
            text = element_text(size = 16))

@@

``` {r 16.2.2. rmANOVA, echo=FALSE,fig.align="center",fig.width=5,fig.height=4, results = 'asis'}
temp <- dataAll %>%
        filter(Choice < 2) %>%
        group_by(SubjID, Cost, Handling, Reward) %>%
        summarise(pAccept = mean(Choice))

## REPEATED MEASURES ANOVA
# I think I like this the most
prop.aov <- with(temp, aov(pAccept ~ factor(Cost) * Handling * Reward + 
                                 Error(SubjID / (Handling * Reward))))

@@

``` {r 16.2.2. rmANOVA plot, echo = FALSE, fig.align = "center", fig.width = 10, fig.height = 6}
# prep data to plot the p(accept) based on reward, handling time and cost type
trueProp <- dataAll %>%
        filter(Choice < 2) %>%
        group_by(SubjID, Cost, Handling, Reward, optimal) %>%
        summarise(pAccept = mean(Choice)) %>%
        group_by(Cost, Handling, Reward, optimal) %>%
        summarise(meanComplete = mean(pAccept),
                  SE = sd(pAccept) / sqrt(length(pAccept)))

@@

``` {r 16.3.1. RTs, echo=FALSE,fig.align="center",fig.width=6,fig.height=4}
# Proportion of forced travels per subject and cost type
propFails <- dataAll %>% 
  filter(Cost != "Wait") %>% 
  group_by(SubjID, Cost) %>% 
  summarise(propFails = mean(rawChoice == 2)) %>%
  group_by(Cost) %>% 
  summarise(meanFT = mean(propFails))

# Reaction times ECDFs including offer and in-trial quits
(RTs <- dataAll %>%
  #mutate(RT = ifelse(Choice != 0, RT + 2, RT)) %>%
  filter(rawChoice < 2) %>% # change this to 1 to reproduce the original plot, since it would only plot the cdf of quits
  ggplot(aes(RT, color = Cost)) + 
    #stat_ecdf(lwd = 1.2, geom = "smooth") +
    geom_line(aes(y = 1 - ..y..), stat = 'ecdf', lwd = 1.5) +
    geom_vline(xintercept = 1, lty = 2) + # end of the grace period for physical trials
    scale_color_manual(values = cols) +
    scale_x_continuous(breaks = seq(14)) +
    labs(x = "Reaction Time", y = "Proportion Accepted") +
    theme(panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(colour = "black"),
      text = element_text(size = 16)))
